Grove: A Convergent Collaborative Structure-Editor Calculus

### 2021-01-14 grove-paper.tex
section 3 updates with Cyrus

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{lrlll}
      \textrm{Patterns:}          & q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \{q | \ldots | q\} \\
      \textrm{Types:}             & \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \{ \tau | \ldots | \tau \} \\
      \textrm{Expressions:\qquad} & e    & {}\in Exp & {}::={} & x^u \mid \lambda^u q \mathord{:}\tau.e \mid (e~e)^u \mid n^u \\
                                  &      &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \{e|\ldots|e\} \\
    \end{array}
  \]
  \caption{Syntax as a grammar for the lambda calculus that we are considering.}
  % TODO add references into multiparent store and cycles store
  \label{fig:Syntax as a grammar}
\end{figure}

@@

The language we are modeling is shown in \autoref{fig:Syntax as a grammar},
standard for a simply-typed lambda calculus except that all sorts have empty
holes (represented by $\hole$) and conflict holes (of the form $\{q | \ldots |
q\}$). In particular, the argument to a lambda can be either a bound variable or
a hole. All constructors carry a superscripted identifier $u \in \U$.

% In order to model this we define

% Turn that into a constructor set K + position set P + arity function (has type: K -> $\wp$(P)).

@@

% NOTE base grammar has holes. rooted expression grammar has refs, conflicts, and multiparents
% NOTE ref = expression reference or vertex reference

@@

Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$,
and connects the child position $p$ of source vertex $v$ to target vertex $v^\prime$.
The state of each edge, $\e$, is determined by the corresponding edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$.
When $G(\e) = \bot$, $\e$ has not been created yet.
When $G(\e)$ is $\Plus$, the edge has been created.
When $G(\e)$ is $\Minus$, the edge has been destroyed.
The total ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.

@@

@@

declarative section: give a hint what decomposition does (incl. wreaths): take a
graph, turn it into a grove (i.e., 4-tuple) which is what we actually present to
users and use for things like type checking. In this section, we present an
algorithm that declaratively say what a grove is and how it connects to the
graph.

We can't turn the graph into a single tree, so we turn it into a forest. A
forest is a collection of trees, in fact it's a 4-tuple consisting of a rooted tree
and 3 sets of trees: mp, d, sc. Before discussing what these sets are, let's 
define what a tree is.

>>>>

\begin{figure}
  \[
    \arraycolsep=0pt
    \begin{array}{lrlll}
      q    & {}\in Pat & {}::={} & x^u \mid \hole \mid \{q | \ldots | q\} \mid \#u \\
      \tau & {}\in Typ & {}::={} & \tau \rightarrow^u \tau \mid Num^u \mid \hole \mid \{ \tau | \ldots | \tau \} \mid \#u \\
      e    & {}\in Exp & {}::={} & x^u \mid \lambda^u q \mathord{:}\tau.e \mid (e~e)^u \mid n^u \\
           &           &         & ~\mid e~\texttt{+}^u~e \mid e~\texttt{*}^u~e \mid \hole \mid \{e|\ldots|e\} \mid \#u \\
    \end{array}
  \]
  \caption{Base syntax, as a grammar, for the lambda calculus that we are considering.}
  \label{fig:Base syntax}
\end{figure}

@@

The language we are modeling is shown in \autoref{fig:Base syntax}, standard for
a simply-typed lambda calculus except that all sorts have empty holes
(represented by $\hole$) and conflict holes (of the form $\{q | \ldots | q\}$).
In particular, the argument of a lambda can be either a bound variable or a
hole. Additionally, all non-hole forms carry a superscripted identifier $u \in
\U$.

% Turn that into a constructor set K + position set P + arity function (has
% type: K -> $\wp$(P)).

@@

@@

Each edge $\e = (u, v, p, v^\prime)$ has a unique name, $u \in \U$, and connects
the child position $p$ of originating vertex $v$ to destination vertex
$v^\prime$. The state of each edge, $\e$, is determined by the corresponding
edge state, $G(\e) \in \Sigma = \{ \bot, \Plus, \Minus \}$. When $G(\e) = \bot$,
$\e$ has not been created yet. When $G(\e)$ is $\Plus$, the edge has been
created. When $G(\e)$ is $\Minus$, the edge has been destroyed. The total
ordering $\bot \sqsubset \Plus \sqsubset \Minus$ forms a lattice over $\Sigma$.

@@

% TODO: clean up and move rooted expression grammar

% NOTE base grammar has holes. rooted expression grammar has refs, conflicts,
% and multiparents

% TODO add references into multiparent store and cycles store

@@

% NOTE: this whole section is for describing the 4-tuple

Tree-like structures are often easier to work with than arbitrary graphs
directly, for example to present the graph to users, or for type checking, so we
decompose it into a grove.

A grove is a 4-tuple of the following form:
\[
  (e, MP, NP, W)
\]
The first component is an expression, governed by the grammar in
\autoref{fig:Base syntax}, determined by starting at the root of the graph and
including every connected vertex that has a single parent. In positions where
there are no children, we leave an empty hole, $\hole$. In positions where there
are multiple children, we leave a conflict hole of the form $\{e_1 | \ldots |
e_n\}$. Whenever we encounter a vertex with multiple parents, we refer to it
indirectly by reference $\#u$. We collect the expressions corresponding to
vertices with multiple parents in the set $MP$.

These first two components do not account for vertices with no parents. We
include the expressions corresponding to these vertices in the set $NP$.

We've accounted for vertices with multiple parents, with no parents, and with
one parent connected to the root. This leaves only vertices with one parent that
are not connected to the root. These vertices form structures we call wreaths.

Wreaths:

Three components are  <describe each of the components:
The first component is ... trees. The fourth component is ... wreaths>.

A forest is a collection of trees, in fact it is a 4-tuple consisting of
a rooted tree $R$ and 3 sets of trees: $MP, D, W$. Before discussing what these
sets are, let's define what a tree is.
